---
title: "Evaluation of impact of various factors(specially transmission) on MPG in Cars"
author: "Ritesh Kumar Malaiya"
date: "Saturday, July 26, 2015"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
---

# Evaluate Dataset

```{r warning =FALSE, echo=FALSE}
library(datasets)
data(mtcars)
```

## FIT MODEL

1) Identify variables having Linear relationship with MPG

```{r warning =FALSE, echo=FALSE}
mtcars$am[mtcars$am == 0] = "Automatic"
mtcars$am[mtcars$am == 1] = "Manual"

mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$vs <- factor(mtcars$vs)
mtcars$cyl <- factor(mtcars$cyl)

par(mfrow=c(2,5))

mpg <- mtcars$mpg
variables <- colnames(mtcars)
variables <- variables[variables != 'mpg']


for (i in 1: length(variables)) {
  x <- variables[i]
  plot(mpg~mtcars[,x],xlab = x, ylab = "Mileage Per Gallon (MPG)")
  abline(lm(mpg~mtcars[,x]))
}
```

Based on above graph, we can observe that **cyl, hp, drat, wt, vs, am, disp, carb, qsec** variables seems to have a better linear relationship with mpg compared to others. 


### Identify Corelation

Let's try to observe the correlation between these variables (except cyl, am and vs, since these are categorical variable)
```{r echo=FALSE}
variables <- c('hp', 'drat', 'wt', 'disp', 'carb', 'qsec')

for (i in 1:length(variables)) {
 mtcars[,i] <- as.numeric(mtcars[,i])  
}

sapply(variables,function(x) {
  cor(y = mpg, x = mtcars[,x], use = "complete")
})

```

At this point we have a fair idea how variables are correlated to mpg. (closer the correlation value to -1 or 1, better is the correlation)

### Fit a Linear Model

Here since, transmission is a factor variable we cannot efficiently fit linear model for mpg based on transmission only. Hence based on correlation score let's try to fit an initial model using cyl, hp, wt


```{r message=FALSE, results='hide'}
library(MASS)
lm_initial <- lm(mpg ~ am + hp + drat + wt + disp + carb + qsec, data = mtcars)
step <- stepAIC(lm_initial, direction="both")
```

Model recommendation by AIC

```{r results='markup'}
step$anova # display results
```

Let's cross check the model using Leaps package

```{r warning=FALSE, message=FALSE, results='hide' }
library(leaps)
```
```{r}
leaps <- regsubsets(mpg ~ am + hp + drat + wt + disp + carb + qsec, data = mtcars, nbest = 3)
par(mfrow=c(1,1))
plot(leaps,scale="r2")

```

From above analysis, we can conclude that below should be the final model
```{r}

lm <- lm(mpg ~ am + wt + qsec, data = mtcars)
```

```{r echo=FALSE}
par(mfrow=c(2,2))
summary(lm)
plot(lm)
```

### Check Relative Importance of the Model:
```{r message = FALSE, warning = FALSE, results='hide'}

library(relaimpo)

calc.relimp(lm,type=c("lmg","last","first","pratt"), rela=TRUE)
```

```{r message = FALSE, results='hide'}
boot <- boot.relimp(lm, b = 1000, type = c("lmg", "last", "first", "pratt"), rank = TRUE, diff = TRUE, rela = TRUE)
# booteval.relimp(boot) # print result
```

```{r}
plot(booteval.relimp(boot,sort=TRUE)) # plot result
```

we can also verify our model using chi square test whose value = ***`r 1-pchisq(2.306,26)`***. To cross check whether this can be a case of overfitting. Let's try to drop a variable

```{r}

drop1(lm, test="Chisq")

```
As we can see, we do not require to drop any variable.


## Coefficients as per the model

I guess now we can safely assume that below are the coef of this model

```{r}

lm$coef

```


# Questions of Interest

- **Is an automatic or manual transmission better for MPG**
From above coefficients we can understand that ***Manual Transmission is better*** than Automatic Transmission.


- **Quantify the MPG difference between automatic and manual transmissions**
If we keep weight and qsec constant (mean) then extra milega that will be provided by Manual Transmission as compared to Automatic is `r lm$coef[2]`. However, for every increase in weight by single unit gain in mileage is `r lm$coef[1] + lm$coef[2] + lm$coef[3]` and for every increase in qsec gain in mileage is `r lm$coef[1] + lm$coef[2] + lm$coef[4]` more than Automatic Transmission.

# Executive Summary
For a Car having same weigth Manual Transamission gives better mileage than Automatic Transmission. Also, for every increase in weight and / or qsec, corresponding increase in Manual transmission is in range of > `r lm$coef[2]` and < `r lm$coef[1] + lm$coef[2] + lm$coef[3]  + lm$coef[4]`

