---
title: "Impact of various factors(specially transmission) on MPG in Cars"
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---
# Executive Summary
For a Car having same weigth, Manual gives better mileage than Automatic. Also, for every increase in weight and / or qsec, corresponding increase in Manual is in range of > 2.9358 and < 9.863.

# Evaluate Dataset
```{r warning =FALSE, echo=FALSE}
library(datasets)
data(mtcars)
```
## FIT MODEL
### Identify Corelation
Let's try to observe the correlation between mpg and other variables (except cyl, am and vs, since these are categorical variable)
```{r echo=FALSE}
mtcars$am[mtcars$am == 0] = "Automatic"
mtcars$am[mtcars$am == 1] = "Manual"

mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$vs <- factor(mtcars$vs)
mtcars$cyl <- factor(mtcars$cyl)

mpg <- mtcars$mpg

variables <- c('hp', 'drat', 'wt', 'disp', 'carb', 'qsec')

for (i in 1:length(variables)) {
 mtcars[,i] <- as.numeric(mtcars[,i])  
}

sapply(variables,function(x) {
  cor(y = mpg, x = mtcars[,x], use = "complete")
})

```
### Fit a Linear Model
Based on correlation found above let's use AIC to choose attributes for Linear Model
```{r message=FALSE, results='hide'}
library(MASS)
lm_initial <- lm(mpg ~ am + hp + drat + wt + disp + carb + qsec, data = mtcars)
step <- stepAIC(lm_initial, direction="both")
```
Model recommendation by AIC
```{r results='markup'}
step$anova # display results
```
From above analysis, we can conclude that below should be the final model
```{r}
lm <- lm(mpg ~ am + wt + qsec, data = mtcars)
```
## Plot of Model
```{r echo=FALSE}
par(mfrow=c(2,2))

plot(lm)
```
## Coefficients as per the model
```{r echo=FALSE}

summary(lm$coef)
```
# Questions of Interest
- **Is an automatic or manual transmission better for MPG**
From above coefficients we can understand that ***Manual Transmission is better*** than Automatic Transmission.
- **Quantify the MPG difference between automatic and manual transmissions**
If we keep weight and qsec constant (mean) then extra milega that will be provided by Manual as compared to Automatic is `r lm$coef[2]`. However, for every increase in weight by single unit gain in mileage is `r lm$coef[1] + lm$coef[2] + lm$coef[3]` and for every increase in qsec gain in mileage is `r lm$coef[1] + lm$coef[2] + lm$coef[4]` more than Automatic.

