---
title: "Human Activity Recognition"
author: "Ritesh Kumar Malaiya"
date: "August 18, 2015"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(caret, quietly = T, warn.conflicts = F)
library(e1071, quietly = T, warn.conflicts = F)
if(!require(RANN)) {
  install.packages("RANN")
  require(RANN)
}
if(!require(doParallel)) {
  install.packages("doParallel")
  require(doParallel)
}
if(!require(klaR)) {
  install.packages("klaR")
  require(klaR)
}
if(!require(MASS)) {
  install.packages("MASS")
  require(MASS)
}
```


# Read Data

Let's read the csv training data - pml-training.csv. This will be sub divided into training and test sets using cross validation.After a model is trained and tested on this dataset we will apply the model on the pml-testing.csv file provided.


```{r}
pml <- read.csv("pml-training.csv")
pmlTestFinal <- read.csv("pml-testing.csv")
pmlTrain <- pml

```

# Exploratory Data Analysis

## Check Dimension of Data
```{r}
dim(pmlTrain)
```

## Check all possible classes
```{r}
unique(pmlTrain$classe)
```

## Variables format conversion

1. Let's find out variables which may have been wrongly identified as factor/character but are actually numeric variables
```{r}
pmlTrain_class <- sapply(pmlTrain, class)
non_integer_columns <- names(pmlTrain[, which(pmlTrain_class != 'integer' & pmlTrain_class != 'numeric')])
non_integer_columns
```

* Let's check the data present for these columns
```{r}
head(pmlTrain[complete.cases(pmlTrain[, non_integer_columns]), non_integer_columns])

```
* First look in the name list we can identify that num_window, raw_timestamp_part_1, raw_timestamp_part_2, user_name,cvtd_timestamp, new_window, classe variables cannot be converted to numeric data
```{r}
pmlTrain <- pmlTrain[, !names(pmlTrain) %in% c("X","user_name", "new_window" ,"classe", "cvtd_timestamp", "raw_timestamp_part_1", "raw_timestamp_part_2", "num_window") ]
```

* For rest variables let's convert them to numeric variables

```{r}
pmlTrain <- data.frame(sapply(pmlTrain, as.numeric)) 
pmlTrain$new_window <- pml$new_window

```



# Preprocessing of Data

## Column-wise Preprocessing

### Dummy Data
Per data analysis done above, we now know that new_window variable is a factor variable and need preprocessing.

```{r}

d <- dummyVars(~new_window, data = pmlTrain)
temp <- data.frame(predict(d, pmlTrain))

pmlTrain$new_window.no <- temp$new_window.no
pmlTrain$new_window.yes <- temp$new_window.yes

pmlTrain <- pmlTrain[, !names(pmlTrain) %in% c("new_window") ]

```

### Near Zero Variables
```{r}
nzr <- nearZeroVar(pmlTrain, saveMetrics = T)
```

Per Near Zero Variance test, we should filter below features from the data
```{r}
row.names(nzr[nzr$nzv == T,])
nzr <- nearZeroVar(pmlTrain)
pmlTrain <- pmlTrain[,-nzr]
```

### Find Corelation
```{r}
highlyCor <- findCorrelation(pmlTrain, exact = F)
pmlTrain <- pmlTrain[ , - highlyCor[complete.cases(highlyCor)]]

```


## Row-wise Preprocessing
### Impute
```{r}
preObj <- preProcess(x = pmlTrain)

preObj <- preProcess(x = pmlTrain, method = c("knnImpute"))
pmlTrain_rel <- data.frame(predict(preObj, pmlTrain))
pmlTestFinal_rel <- data.frame(predict(preObj, pmlTestFinal_rel))

```

### Principal Component Analysis
```{r}
preObj <- preProcess(x = pmlTrain, method = c("pca"))
pmlTrain <- data.frame(predict(preObj, pmlTrain))

```

### Split Data into Training & Test Set
```{r}

pmlTrain$classe <- pml$classe
pmlTrain <- pmlTrain[, !names(pmlTrain) %in% c("classe")]


inTrain <- createDataPartition(y = pmlTrain$classe, p = .7, list = F)
pmlTrain_rel<- pmlTrain[inTrain,]
pmlTest<- pmlTrain[-inTrain,]

```


# Train a model

## Define Train control
```{r}
registerDoParallel(cores=4)
control <- trainControl(method="repeatedcv", number=2, repeats=2)
```

## Train and Plot a model
```{r cache=TRUE}
model <- train(classe~., data=pmlTrain_rel, method="gbm", trControl=control)
plot(model)
```

## Prediction
```{r}
pmlTrain_predict <- predict(model, newdata = pmlTest[, !names(pmlTrain) %in% c("classe")])
pmlTrain_predict <- predict(model)
cnfM <- confusionMatrix(pmlTest$classe, pmlTrain_predict)
cnfM
pmlTrain_predict
```
Based on the confusion matrix above, we can see that the Accuracy of model is 55%.


```{r}

pml_test <- read.csv("pml-testing.csv")

pml_test <- pml_test[, !names(pml_test) %in% c("X","user_name", "new_window" ,"classe", "cvtd_timestamp", "raw_timestamp_part_1", "raw_timestamp_part_2", "num_window") ]
pml_test <- data.frame(sapply(pml_test, as.numeric)) 

pmlTrain_predict <- predict(model, newdata = pml_test[, !names(pml_test) %in% c("classe")])

pmlTrain_predict
```
