---
title: "Human Activity Recognition"
author: "Ritesh Kumar Malaiya"
date: "August 18, 2015"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(caret, quietly = T, warn.conflicts = F)
library(e1071, quietly = T, warn.conflicts = F)
if(!require(RANN)) {
  install.packages("RANN")
  require(RANN)
}
if(!require(doParallel)) {
  install.packages("doParallel")
  require(doParallel)
}
if(!require(klaR)) {
  install.packages("klaR")
  require(klaR)
}
if(!require(MASS)) {
  install.packages("MASS")
  require(MASS)
}
```


# Read Data

Let's read the csv training data - pml-training.csv. This will be sub divided into training and test sets using cross validation.After a model is trained and tested on this dataset we will apply the model on the pml-testing.csv file provided.


```{r}
pml <- read.csv("pml-training.csv")
pmlTestFinal <- read.csv("pml-testing.csv")

pmlTrain <- pml

```

# Exploratory Data Analysis

## Check Dimension of Data
```{r}
dim(pmlTrain)
```

## Check all possible classes
```{r}
unique(pmlTrain$classe)
```

## Variables format conversion

1. Let's find out variables which may have been wrongly identified as factor/character but are actually numeric variables
```{r}
pmlTrain_class <- sapply(pmlTrain, class)
non_integer_columns <- names(pmlTrain[, which(pmlTrain_class != 'integer' & pmlTrain_class != 'numeric')])
non_integer_columns


pmlTestFinal_rel <- pmlTestFinal[, !names(pmlTestFinal) %in% c("X","user_name", "new_window" ,"classe", "cvtd_timestamp", "raw_timestamp_part_1", "raw_timestamp_part_2", "num_window") ]



```

* Let's check the data present for these columns
```{r}
head(pmlTrain[complete.cases(pmlTrain[, non_integer_columns]), non_integer_columns])

```
* First look in the name list we can identify that num_window, raw_timestamp_part_1, raw_timestamp_part_2, user_name,cvtd_timestamp, new_window, classe variables cannot be converted to numeric data
```{r}
pmlTrain <- pmlTrain[, !names(pmlTrain) %in% c("X","user_name", "new_window" ,"classe", "cvtd_timestamp", "raw_timestamp_part_1", "raw_timestamp_part_2", "num_window") ]
```

* For rest variables let's convert them to numeric variables

```{r}
pmlTrain <- data.frame(sapply(pmlTrain, as.numeric)) 
pmlTrain$classe <- pml$classe

pmlTestFinal_rel <- data.frame(sapply(pmlTestFinal_rel, as.numeric)) 

sum(is.infinite(pmlTrain))
sum(is.nan(pmlTrain))

inTrain <- createDataPartition(y = pmlTrain$classe, p = .7, list = F)
pmlTrain_rel<- pmlTrain[inTrain,]
pmlTest<- pmlTrain[-inTrain,]

library(dplyr)



preObj <- preProcess(x = select(pmlTrain_rel, -classe) , method = c("center", "scale", "knnImpute"))

pmlTrain_rel <- data.frame(predict(preObj, pmlTrain_rel))

pmlTest <- data.frame(predict(preObj, pmlTest))
pmlTestFinal_rel <- data.frame(predict(preObj, pmlTestFinal_rel))


```

