---
title: "Practical Machine Learning - Preprocessing"
author: "Ritesh Kumar Malaiya"
date: "August 16, 2015"
output: html_document
---

```{r}

library(caret)
library(kernlab)
library(dplyr)

data(spam)

inTrain <- createDataPartition( y = spam$type, p = .75, list = F)

training <- spam[inTrain,]
testing <- spam[-inTrain,]

```

Example of Variable Skeweness

```{r}

hist(training$capitalAve, main = "", xlab = "ave. capital run length")

```

Let's use Preprocess feature of Caret Package. In below example 58th column in the actual outcome.

```{r}

preObj <- preProcess(x = training[,-58], method = c("center", "scale"))

str(preObj)

```

BoxCox
```{r}
#install.packages("e1071")
library(e1071)
preObj <- preProcess(x = training[,-58], method = c("BoxCox"))
str(preObj)
```

Impute
```{r}
preObj <- preProcess(x = training[,-58], method = c("knnImpute"))
str(preObj)
``` 

Preprocessing with Principal Components
Outliers can be bad for pca.


```{r}
typeColor <- (spam$type == "spam")*1+1
prComp <- prcomp(log10(spam[,-58] + 1))


```


Regression using Multiple Covariates

```{r}
library(caret)
library(ISLR)
library(dplyr)


data(Wage)

inTrain <- createDataPartition(y=Wage$wage, p = .75, list = F)

training <- Wage[inTrain,]
testing <- Wage[-inTrain,]

qplot(age, wage, colour=jobclass, data=training)

qplot(age, wage, colour=education, data=training)

modFit <- train(wage~age + jobclass + education, method = "lm", data=training)

modFit$finalModel

plot(modFit$finalModel,1)
```

Color by variables not used in the model

```{r}
qplot(modFit$finalModel$fitted.values,modFit$finalModel$residuals, colour = race, data=training)

```

Plot by index, means row numbers. Below means some variable is missing from model. This is because there is a pattern visible based on index

```{r}

plot(modFit$finalModel$residuals, pch = 19)

```

Plot predicted vs Actuals

```{r}

pred <- predict(modFit, testing)
qplot(wage, pred, colour=year, data = testing)

```
